---
title: "PA1_template"
author: "Narayan Menon"
date: "December 8, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library}
library(dplyr)
```

Load the file from working directory

```{r}
act <- read.csv("activity.csv")

#convert the date from factor
act$date <- as.Date(act$date)
```


Calculate total steps in a day
```{r}
daily_act <- aggregate(act$steps,by=list(act$date),sum)
```

Draw a histogram of total steps in a day
```{r}
hist(daily_act$x)
```

Mean number of steps in a day
```{r}
mean(daily_act$x,na.rm=TRUE)
```

Median number of steps in a day
```{r}
median(daily_act$x,na.rm=TRUE)
```

Calculate total steps by interval across all days
```{r}
daily_interval_act <- aggregate(act$steps,by=list(act$interval),mean,na.rm=TRUE)
```

Graph showing average number of steps by intervals
```{r}
plot(x=daily_interval_act$Group.1,y=daily_interval_act$x,type="l",xaxt="n",
     xlab="Intervals",ylab="Average steps")
axis(1, at=seq(0, 2355, by=50))
```


Interval that has highest average number of steps
```{r}
daily_interval_act[daily_interval_act$x==max(daily_interval_act$x),]$Group.1
```

Count of missing values in the dataset
```{r}
sum(is.na(act$steps))
```

Function to return mean steps of an interval
```{r}
getIntMean <- function(interval) {
    daily_interval_act[daily_interval_act$Group.1==
                         as.numeric(interval),][[2]]}
```

Function to get adjusted steps to remove all NAs
```{r}
getAdjustedSteps <- function(steps,intMean) {
    if (is.na(steps)) {
        intMean
    }
    else {
        steps
    }
}
```

Create a new data frame that includes a column that has no NAs for steps
```{r}
act1 <- act %>% rowwise() %>% mutate(intMean=getIntMean(interval))
act2 <- act1 %>% rowwise() %>% mutate(adjSteps=getAdjustedSteps(steps,intMean))
```


Histogram of total steps each day
```{r}
daily_adj_act <- aggregate(act2$adjSteps,by=list(act2$date),sum)
hist(daily_adj_act$x)
```

Adjusted mean
```{r}
mean(daily_adj_act$x)
```
Adjusted median
```{r}
median(daily_adj_act$x)
```
Add a factor variable to indicate if it is weekday or weekend
```{r}
act2$dayInd <- factor(weekdays(act2$date) %in% c('Saturday','Sunday'),
                      levels=c('TRUE','FALSE'), labels=c('weekend','weekday'))
```

Total steps by interval across all weekdays & weekends 
```{r}
act2weekday <- act2 %>% filter(dayInd == 'weekday')
act2weekend <- act2 %>% filter(dayInd == 'weekend')
daily_adj_weekday_interval_act <- aggregate(act2weekday$adjSteps,by=list(act2weekday$interval),mean)
daily_adj_weekend_interval_act <- aggregate(act2weekend$adjSteps,by=list(act2weekend$interval),mean)
```

Graph of average steps for weekdays and weekends
```{r}
plot(x=daily_adj_weekday_interval_act$Group.1,y=daily_adj_weekday_interval_act$x,type="l",xaxt="n",
      xlab="Intervals",ylab="Average steps",col="blue")
points(x=daily_adj_weekend_interval_act$Group.1,y=daily_adj_weekend_interval_act$x,type="l",col="red")
axis(1, at=seq(0, 2355, by=50))
legend("topright", legend=c("Weekend", "Weekday"),
      col=c("red", "blue"), lty=1, cex=0.8)
```

